---
title: "Instacart Exploratory Analysis"
output:
  html_document: default
  html_notebook: default
---


```{r}
source("readDataToMemory.R")
readInstacart()

library(DBI)
library(ggplot2)
library(ggthemes)

src_tbls(sc)
```

### Order products

`order_products__*.csv`

These files specify which products were purchased in each order.

```{r}
dbGetQuery(sc, 
"SELECT * 
FROM order_products__prior_tbl 
LIMIT 10")
```
 
## orders

`orders.csv`

This file tells to which set (prior, train, test) an order belongs. 


```{r}
orders_prev <- dbGetQuery(sc, 
"SELECT * 
FROM orders_tbl 
LIMIT 100")
orders_prev
```
 
```{r}
dbGetQuery(
  sc,
  "
SELECT COUNT(DISTINCT product_id) AS products
FROM order_products__prior_tbl
  "
  )
```

```{r}
products_popularity_txt <- "
SELECT product_id
,   COUNT(1) AS n_orders
FROM order_products__prior_tbl
GROUP BY product_id
ORDER BY n_orders DESC
LIMIT 50
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

## Products

`products.csv`

```{r}
dbGetQuery(sc, "SELECT * FROM products_tbl LIMIT 20")
```

## Joins

```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

```{r}
products_popularity %>%
  ggplot(
    aes(reorder(product_name, n_orders, function(x) x), 
        n_orders)) +
  geom_bar(stat="identity", fill='darkorange') +
  coord_flip() + 
  scale_y_continuous(label=scales::comma) +
  xlab("product") +
  ylab("Number of orders") +
  theme_minimal()
```

```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

## Questions

* Which day of week people order most fequently?

```{r}
ex1 <- "select order_dow, count(*) as n 
        from orders_tbl 
        group by order_dow
        order by count(*) desc
        LIMIT 1"
dbGetQuery(sc, ex1) #0 -> diumenge
```

* The most popular hour of doing order? 
```{r}
ex2 <- "select order_hour_of_day, count(*) as n 
        from orders_tbl 
        group by order_hour_of_day
        order by count(*) desc
        LIMIT 1"
dbGetQuery(sc, ex2)
```

* The most reordered products?
```{r}
ex3 <- "select product_id, sum(reordered)
        from order_products__prior_tbl
        group by product_id
        order by sum(reordered) desc
        LIMIT 5"
dbGetQuery(sc, ex3)

```

* The less reorderd product?
```{r}
ex4 <- "select product_id, sum(reordered)
        from order_products__prior_tbl
        group by product_id
        order by sum(reordered)
        LIMIT 10"
dbGetQuery(sc, ex4)
```

* Plot histogram of number of products in an order.
```{r}
ex5 <- "select order_id, count(*)
        from order_products__prior_tbl
        group by order_id"
data <- dbGetQuery(sc, ex5) 
```

```{r}
dbGetQuery(sc, 
"
SELECT n_products,
      COUNT(order_id) AS n_orders
FROM ( 
      SELECT order_id,
          COUNT(1) AS n_products
      FROM order_products__prior_tbl
      GROUP BY order_id) op
GROUP BY n_products
")%>%
  ggplot(aes(n_products, n_orders)) + 
  geom_bar(stat = "identity") +
  theme_minimal()
```

```{r}

```




```{r}
dbGetQuery(sc, 
"SELECT order_id
,   COUNT(1) AS n_products
FROM order_products__prior_tbl
GROUP BY order_id
LIMIT 100
")
```

* What is the most popular department? 

```{r}
dbGetQuery(sc, "DESC orders_tbl")
```


## Whisker, magrittr

```{r}
library(whisker)
library(magrittr)

product_times_ordered_txt <- "
SELECT n_orders
,   COUNT(DISTINCT user_id) AS n_users
FROM (
    SELECT o.user_id
    ,   COUNT(DISTINCT op.order_id) AS n_orders
    FROM (
        SELECT order_id
        FROM order_products__prior_tbl
        WHERE product_id = {{productId}}
    ) op
    LEFT JOIN (
        SELECT  user_id
        ,   order_id
        FROM orders_tbl) o
    ON op.order_id = o.order_id
    GROUP BY o.user_id) opo
GROUP BY n_orders
ORDER BY n_orders
"

data <- list(productId = 24852)

product_times_ordered_txt %>%
  whisker.render(data) %>%
  cat
```
```{r}
product_times_ordered_txt %>%
  whisker.render(data) %>%
  dbGetQuery(sc, .) %>%
  ggplot(aes(n_orders, n_users)) + 
  geom_line() +
  theme_minimal()
```

## Questions

* For given department, given a list of 10 the most popular product (with templating department).
```{r}
dbGetQuery(sc, 
"SELECT *
FROM products_tbl
limit 20") 
```

* For given day of the week, rank department popularity.
```{r}

```


## More on orders (Window Functions)



```{r}
dbGetQuery(sc, 
"
SELECT days_since_prior_order
,   COUNT(1) AS N
FROM orders_tbl
GROUP BY days_since_prior_order
") 
```


```{r}
dbGetQuery(sc, 
"SELECT * 
FROM orders_tbl
WHERE user_id <= 10
")
```

Per a casa usuari (agafem només usuaris fins al 10), agafem la primera ordre que va fer i la guardem a la columna first_order.
```{r}
dbGetQuery(sc, 
"SELECT user_id
,   order_id
,   FIRST_VALUE(order_id)  OVER(PARTITION BY user_id ORDER BY order_number) AS first_order
FROM orders_tbl
WHERE user_id <= 10 
")
```
Perquè surti només una línia per usuari:
```{r}
dbGetQuery(sc, 
"SELECT distinct user_id, FIRST_VALUE(order_id) OVER(partition by user_id ORDER BY order_number) AS first_order
FROM orders_tbl
WHERE user_id <= 10 
ORDER BY user_id
")
```


Ara volem buscar el producte més comú en les cistelles grans:
* Product that is usually orders in numerous baskets

```{r}
"
SELECT product_id
,   AVG(products_number) AS avg_products_number
FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn
GROUP BY product_id
ORDER BY avg_products_number DESC
LIMIT 100
" %>%
dbGetQuery(sc, .)
```
```{r}
"
SELECT product_id
,   AVG(products_number) AS avg_products_number
FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn
GROUP BY product_id
ORDER BY avg_products_number DESC
LIMIT 100
" %>%
dbGetQuery(sc, .)
```

* Could modify the query and add prduct name?

```{r}
"
SELECT pn.product_id, product_name
,   AVG(products_number) AS avg_products_number
FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn LEFT JOIN PRODUCTS_TBL t2 on pn.product_id=t2.product_id
GROUP BY pn.product_id, product_name
ORDER BY avg_products_number DESC
LIMIT 20
" %>%
dbGetQuery(sc, .)
```


* Product that is usually orders in small baskets.

* For 10 first users and all their orders find days since first order.

```{r}
orders %>%
  filter(user_id <= 10) %>%
  select(user_id, order_number, days_since_prior_order) %>%
  arrange(user_id, order_number) %>%
  mutate(days_since_first_order = cumsum(days_since_prior_order)) %>%
  mutate(days_since_first_order=ifelse(is.null(days_since_prior_order), 0, days_since_first_order))
```

```{r}
"SELECT user_id, 
       days_since_prior_order,
       COALESCE(
            SUM(days_since_prior_order) OVER
            (PARTITION BY user_id ORDER BY order_number ROWS BETWEEN UNBOUNDED PRECEDING AND CURRENT ROW),
      e) AS days_since_prior_order,
      order_number
FROM orders_tbl
WHERE user id <= 10
ORDER BY user_id,
    order_number"%>%
 dbGetQuery(sc, .) 
```









